\documentclass[12pt]{article}

\usepackage{blindtext}
\usepackage{pgfplots}
\usepackage{amsmath}
\usepackage{geometry}[paper=letterpaper]

\usepackage{booktabs}
\usepackage{xcolor}
\usepackage[T1]{fontenc}

\let\oldref\ref
\renewcommand{\ref}[1]{(\oldref{#1})} % references in parathesis
\renewcommand*\arraystretch{1.3} % extra vertical space in tables


\newcommand{\dtoprule}{\specialrule{1pt}{0pt}{1.5pt}%
            \specialrule{0.3pt}{0pt}{\belowrulesep}%
            }
\newcommand{\dbottomrule}{\specialrule{0.3pt}{2pt}{1.5pt}%
            \specialrule{1pt}{0pt}{\belowrulesep}%
            }



% All styles and package and commands are defined here.
% This allow to have one preambule used in all |/subfile| .
%\usepackage{mystyle}

\title{EE Bits and Pieces}
\author{A. Nonym}



\begin{document}

%\maketitle


\vfil
\eject

\section{Power}

When we say power we usually are talking about \textsc{average power}. For DC circuits it is nothing but $ P = VA$, for AC circuits average is an integral of instanteneous voltage $v(t)$ and $i(t)$:

\begin{equation}\label{eq:100}
    P_{avg} = \frac{1}{T}\int_0^T v(t) i(t) \,dt.
\end{equation}

Also, specifically for RLC circuit it can be shown


\begin{equation}
    P_{avg} = (V_{rms} \times I_{rms}) cos\phi
\end{equation}

Here, $\phi$ is the angle between voltage and current phasor vectors(assumed constant). The product $(V_{rms} \times I_{rms})$ is the easily measurable quantity obtained by any DMM. On the other hand neither $\phi$ nor $P_{avg}$ are easily measurable. 

\medskip
\begin{center}
The term $VA = (V_{rms} \times I_{rms})$ is called \textsc{apparent power}. 
\end{center}

\medskip
\begin{center}
The term $cos\phi$ is called \textsc{power factor(PF).}
\end{center}

\begin{equation}
 P_{avg} = VA \times PF
\end{equation}


% P 
% real energy consumption
% total power of the system

% VA
% real current consumption
% does not add in general



When we talk about {\it power factors}, we are assuming it can be represented by an equivalent RLC circuit(which may not always be realistic). \textsc{If we did know what RLC circuit it was we could calculate power factor directly}:

\begin{equation}
    \phi = tan^{-1} \bigl( \frac{X_{L} - X_{C}}{R} \bigr), PF = cos \phi
\end{equation}

Power factor relates easily "measurable" apparent power $VA$ term to the less tangible $P_{avg}$.

Calculating power factor requires either computing integral from \ref{eq:100} or finding average power by some other means. Then

\begin{equation}
  PF = \frac{P_{avg}}{VA}
\end{equation}



\vfil
\eject


P versus VA.

\begin{center}
{
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{rcc}
\dtoprule
                                      &  VA         & P  \\ \midrule[0.5pt]
\textsc{measuring}                    &  GREAT      & OK \\
\textsc{estimate total system power}  &  NOT GREAT  & GREAT \\
\textsc{pay power bills}              &  NOT GREAT  & GREAT \\
\textsc{estimate heat}                &  NOT GREAT  & GREAT \\
\textsc{estimate device current draw} &  GREAT      & NOT GREAT   \\
\textsc{estimate system current draw} & NOT GREAT   & OK \\
\dbottomrule
\end{tabular}
}
\end{center}

\end{document}
